Refactoring the `nls()` function in R

### 2022-05-28 RefactoringNLS/RefactoringNLS.Rmd
Edits and cleanup 20220528

---
title: "Refactoring the `nls()` function in R"
author: 
   - John C. Nash \thanks{ retired professor, Telfer School of Management, University of Ottawa}
   - Arkajyoti Bhattacharjee \thanks{Department of Mathematics and Statistics, Indian Institute of Technology, Kanpur}
date: "2021-8-20"
output: 
    pdf_document:
        keep_tex: false
        toc: true
bibliography: ../BibSupport/ImproveNLS.bib
link-citations: yes
linkcolor: red
urlcolor: blue
---

@@

This article reports the particular activities of our Google Summer of Code
project "Improvements to `nls()`" that relate to the R code for that function, which
is intended for the estimation of models written as a formula that has at least
one parameter that is "nonlinear", that is, not estimable via solving a set of 
linear equations. A companion document **Variety in Nonlinear Least Squares Codes** 
presents an overview of methods for the problem which takes a much wider view of the
problem of minimizing a function that can be written as a sum of squared
terms.

Our work has not fully addressed all the issues that we would like to see
resolved, but we believe we have made sufficient progress to demonstrate that
there are worthwhile improvements that can be made to the R function `nls()`.
An important overall consideration in our work has been the maintainability of
the code base that supports the `nls()` functionality, as we believe that the
existing code makes maintenance and improvement very difficult.

@@

With such a range of features and a long history, it is not surprising that the code has become untidy 
and overly patched. It is, to
our mind, essentially unmaintainable. Moreover, its underlying methods can and should be improved. Let
us review some of the issues. We will then propose corrective actions, some of which we have carried out.

@@

A previous issue with `nls()` that prevented it from providing parameter estimates for zero-residual
(i.e., perfect fit) data was corrected thanks to suggestions by one of us. The release 
notes for R 4.1.0 state:

 - For nls(), an optional constant scaleOffset may be added to the denominator of the relative offset
   convergence test for cases where the fit of a model is expected to be exact, thanks to a proposal by 
   John Nash. nls(*, trace=TRUE) now also shows the convergence criterion.

Within the previous standard documentation (**man** or ".Rd" file) `nls()` warned

>  **The default settings of nls generally fail on artificial “zero-residual” data problems.**

>  The nls function uses a relative-offset convergence criterion that compares the numerical imprecision at
  the current parameter estimates to the residual sum-of-squares. This performs well on data of the form $$y = f(x, \theta) + eps$$

>  (with var(eps) > 0). It fails to indicate convergence on data of the form $$y = f(x, \theta)$$

>  because the criterion amounts to comparing two components of the round-off error. To avoid a zero-divide in
  computing the convergence testing value, a positive constant `scaleOffset` should be added to the denominator
  sum-of-squares; it is set in control; this does not yet apply to algorithm = "port".

It turns out that this issue can be quite easily resolved. The key "convergence test" -- more properly
a "termination test" for the **program** rather than testing for convergence of the underlying **algorithm** --
is the Relative Offset Convergence Criterion (see @BatesWatts81). This works by projecting the proposed
step in the parameter vector on the gradient and estimating how much the sum of squares loss function 
will decrease. To avoid scale issues, we use the current size of the loss function as a measure and divide
by it. When we have "converged", the estimated
decrease is very small, as usually is its ratio to the sum of squares. However, in some cases we have the 
possibility of an exact fit and the sum of squares is (almost) zero and we get the possibility of a 
zero-divide failure.

The issue is easily resolved by adding a small quantity to the loss function. To preserve legacy behaviour,
in 2021, one of us (J. Nash) proposed that `nls.control()` have an additional parameter `scaleOffset` 
with a default value of zero for legacy behaviour. Setting it to a small number -- 1.0 is 
a reasonable choice --
allows small-residual problems (i.e., near-exact fits) to be dealt with easily. We call this the
**safeguarded relative offset convergence criterion**.

As noted, we are pleased to report that this improvement is now in the R distributed code as of version 4.1.0.

@@

```
rm(list=ls())
t <- -10:10
y <- 100/(1+.1*exp(-0.51*t))
lform<-y~a/(1+b*exp(-c*t))
ldata<-data.frame(t=t, y=y)
plot(t,y)
lstartbad<-c(a=1, b=1, c=1)
lstart2<-c(a=100, b=10, c=1)
nlsr::nlxb(lform, data=ldata, start=lstart2)
nls(lform, data=ldata, start=lstart2, trace=TRUE)
# Fix with scaleOffset
nls(lform, data=ldata, start=lstart2, trace=TRUE, control=list(scaleOffset=1.0))
sessionInfo()
```

@@

```
> rm(list=ls())
> t <- -10:10
> y <- 100/(1+.1*exp(-0.51*t))
> lform<-y~a/(1+b*exp(-c*t))
> ldata<-data.frame(t=t, y=y)
> plot(t,y)
> lstart2<-c(a=100, b=10, c=1)
> nlsr::nlxb(lform, data=ldata, start=lstart2)
nlsr object: x 
residual sumsquares =  1.007e-19  on  21 observations
    after  13    Jacobian and  19 function evaluations
  name            coeff          SE       tstat      pval      gradient    JSingval   
a                    100     2.679e-11  3.732e+12  1.863e-216  -6.425e-11       626.6  
b                    0.1      3.78e-13  2.646e+11  9.125e-196  -3.393e-08       112.3  
c                   0.51       6.9e-13  7.391e+11  8.494e-204   1.503e-08       2.791  
# Note that this has succeeded. The test in nlsr recognizes small residual problems.
> nls(lform, data=ldata, start=lstart2, trace=TRUE)
40346.    (1.08e+00): par = (100 10 1)
11622.    (2.93e+00): par = (101.47 0.49449 0.71685)
5638.0    (1.08e+01): par = (102.23 0.38062 0.52792)
642.08    (1.04e+01): par = (102.16 0.22422 0.41935)
97.712    (1.79e+01): par = (100.7 0.14774 0.45239)
22.250    (1.78e+02): par = (99.803 0.093868 0.50492)
0.025789  (1.33e+03): par = (100.01 0.10017 0.50916)
6.0571e-08 (7.96e+05): par = (100 0.1 0.51)
4.7017e-19 (1.86e+04): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
Error in nls(lform, data = ldata, start = lstart2, trace = TRUE) : 
  number of iterations exceeded maximum of 50

> nls(lform, data=ldata, start=lstart2, trace=TRUE, control=list(scaleOffset=1.0))
40346.      (1.08e+00): par = (100 10 1)
11622.      (2.91e+00): par = (101.47 0.49449 0.71685)
5638.0      (9.23e+00): par = (102.23 0.38062 0.52792)
642.08      (5.17e+00): par = (102.16 0.22422 0.41935)
97.712      (2.31e+00): par = (100.7 0.14774 0.45239)
22.250      (1.11e+00): par = (99.803 0.093868 0.50492)
0.025789    (3.79e-02): par = (100.01 0.10017 0.50916)
6.0571e-08  (5.80e-05): par = (100 0.1 0.51)
4.7017e-19  (1.62e-10): par = (100 0.1 0.51)
Nonlinear regression model
  model: y ~ a/(1 + b * exp(-c * t))
   data: ldata
     a      b      c 
100.00   0.10   0.51 
 residual sum-of-squares: 4.7e-19

Number of iterations to convergence: 8 
Achieved convergence tolerance: 1.62e-10

> sessionInfo()
R version 4.1.0 (2021-05-18)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Linux Mint 20.2
```

@@

The single convergence criterion of `nls()` leaves out some possibilities that could
be useful for some problems. The package `nlsr` (@nlsr-manual) already offers both 
the safeguarded relative offset test (**roffset**) as well as a **small sum of 
squares** test (**smallsstest**) that compares the latest evaluated sum of squared
(weighted) residuals to a very small multiple of the initial sum of squares. The
multiple uses a control setting `offset` which defaults to 100.0 and we compute
the 4th power of the machine epsilon times this offset.

@@

We do note that `nls()` stops after `maxiter` "iterations". However, for almost all
iterative algorithms, the meaning of "iteration" requires careful examination of the
code. Instead, we prefer to record the number of times the residuals or the jacobian have
been computed and put upper limits on these. Our codes exit (terminate) when these limits
are reached. Generally we prefer larger limits than the default `maxiter=50` of `nls()`,
but that may simply reflect our history of dealing with more difficult problems as we are
the tool-makers; users consult us when things go wrong.

@@

gets over 4000 hits that are spread over the years. 
In some cases this is due
to the failure of the simple finite difference approximation of the Jacobian in the
`numericDeriv()` function that is a part of `nls()`. `nlsr` can use analytic
derivatives, and we can import this functionality to the `nls()` code as an
improvement. See below in the section **Jacobian computation**.

However, the more common source of the issue is that the Jacobian is very close
to singular for some values of the model parameters. In such cases we need to find
an alternative algorithm to the Gauss-Newton iteration of `nls()`. The most common
work-around is the Levenberg-Marquardt stabilization (see @Marquardt1963, @Levenberg1944, 
@jn77ima). Versions of this have been implemented in packages `minpack.lm` and `nlsr`.
and we have prepared very preliminary versions of an `nls` replacement that can incorporate a
version of the Levenberg-Marquardt stabilization. (There are some issues of 
integration with other code structures and of complexity of the computations that
suggest we should use a simplified LM stabilization.) Unfortunately, making this code
work across all of the features of `nls()` has not proved straightforward.

<!-- ?? JN: I forget exact details, but I recall that we could not get this to work with -->
<!-- some ways in which call could be set up. May need to revisit. this. -->

@@

`nls()`, with the `numericDeriv()` function, computes the Jacobian as the "gradient"
attribute of the residual vector. This is implemented as a mix of R and C code, but
we have created a rather more compact version entirely in R in this Google Summer of
Code project. See the document **DerivsNLS.pdf**.

Let us consider an example which presents the infamous error message.

```{r code=xfun::read_utf8('SESingGrad.R')}
```

It should be noted that the selfStart models in the `./src/library/stats/R/zzModels.R` file
provide the Jacobian in the "gradient" attribute of the "one-sided" formula that defines each
model, and these Jacobians may be the analytic forms. The `nls()` function, after computing the
"right hand side" or `rhs` of the residual, then checks to see if the "gradient" attribute is
defined, and, if not, uses `numericDeriv` to compute a Jacobian into that attribute. This code
is within the `nlsModel()` or `nlsModel.pliner()` functions.

@@

While the implementation of subset at the level of the call to `nls()` has a certain
attractiveness, it does mean that the programmer of the solver needs to be aware of the
source (and value) of objects such as the data, residuals and Jacobian. By preference, 
we would implement subsetting by means of zero-value weights, with observation counts
(and degrees of freedom) computed via the numbers of non-zero weights.

@@

A succinct description of the issue is given in: https://stats.stackexchange.com/questions/492955/should-i-use-na-omit-or-na-exclude-in-a-linear-model-in-r

>>>>

---
title: "Refactoring the `nls()` function in R"
author: 
   - John C. Nash \thanks{ retired professor, Telfer School of Management, University of Ottawa}
   - Arkajyoti Bhattacharjee \thanks{Department of Mathematics and Statistics, Indian Institute of Technology, Kanpur}
date: "2021-8-20"
output: 
    pdf_document:
        keep_tex: true
        toc: true
bibliography: ../BibSupport/ImproveNLS.bib
link-citations: yes
linkcolor: red
urlcolor: blue
---

@@

This article considers the features and limitations of the R function `nls()`. It
arose from a Google Summer of Code project "Improvements to `nls()`", a tool 
intended for the estimation of models written as a formula that has at least
one parameter that is "nonlinear". That is, the model is not estimable via solving a 
finite set of 
linear equations. A companion document **Variety in Nonlinear Least Squares Codes** 
presents an overview of methods for the problem which takes a much wider view of the
problem of minimizing a function that can be written as a sum of squared
terms, and of the even more general case of nonlinear regression.

An important overall consideration in our work has been the maintainability of
the code base that supports the `nls()` functionality, as we believe that the
existing code makes maintenance and improvement very difficult. Moreover, legacy
applications and examples mean some improvments are blocked. Thus we have not
addressed all the issues raised, but we believe that the discussion and trials
so far are relevant to improving both `nls()` and R generally.

@@

With such a range of features and a long history, it is not surprising that the code has become untidy 
and overly patched. It is, to
our mind, essentially unmaintainable. Moreover, its underlying methods can and should be improved. Let
us review some of the issues. We will then propose corrective actions, some of which we have carried out
either in demonstrations or in the distributed coed.

@@

A previous issue with `nls()` that prevented it from providing parameter estimates for zero-residual
(i.e., perfect fit) data was corrected thanks to suggestions by one of us.

In the manual page for `nls()` in R 4.0.0 there is the warning

  **Do not use `nls` on artificial "zero-residual" data.**

  The `nls` function uses a relative-offset convergence criterion
  that compares the numerical imprecision at the current parameter
  estimates to the residual sum-of-squares.  This performs well on data of
  the form

  $$  y = f(x, \theta) + eps $$

  (with `var(eps) > 0`).  It fails to indicate convergence on data of the form

  $$  y = f(x, \theta)  $$

  because the criterion amounts to comparing two components of the round-off 
  error.

  If you wish to test `nls` on artificial data please add a noise component, 
  as shown in the example below.

This amounts to admitting R cannot solve perfectly well-posed problems. The 
suggestion that one needs to pollute data with errors is one that should be offensive
to good science.

It turns out that this issue can be easily resolved. The key "convergence test" -- more properly
a "termination test" for the **program** rather than for convergence of the underlying **algorithm** --
is the Relative Offset Convergence Criterion (see @BatesWatts81). This works by projecting the proposed
step in the parameter vector on the gradient and estimating how much the sum of squares loss function 
will decrease. This is divide by the current size of the loss function to avoid scale issues.
When we have "converged", the estimated
decrease is very small, as usually is its ratio to the sum of squares. However, with small residuals,
the sum of squares loss function is (almost) zero and we get the possibility of a 
zero-divide failure.

Adding a small quantity to the loss function before dividing avoids trouble. 
In 2021, one of us (J. Nash) proposed that `nls.control()` have an additional parameter `scaleOffset` 
with a default value of zero. Setting it to a small number -- 1.0 is 
a reasonable choice --
allows small-residual problems (i.e., near-exact fits) to be dealt with easily. We call this the
**safeguarded relative offset convergence criterion**. The default value gives the legacy behaviour.
We are pleased to note that this improvement is now in the R distributed code as of version 4.1.0.

@@

```
t <- -10:10
y <- 100/(1+.1*exp(-0.51*t))
lform<-y~a/(1+b*exp(-c*t))
ldata<-data.frame(t=t, y=y)
plot(t,y)
lstartbad<-c(a=1, b=1, c=1)
lstart2<-c(a=100, b=10, c=1)
nlsr::nlxb(lform, data=ldata, start=lstart2)
nls(lform, data=ldata, start=lstart2, trace=TRUE)
# Fix with scaleOffset
nls(lform, data=ldata, start=lstart2, trace=TRUE, control=list(scaleOffset=1.0))
sessionInfo()
```

@@

```
> t <- -10:10
> y <- 100/(1+.1*exp(-0.51*t))
> lform<-y~a/(1+b*exp(-c*t))
> ldata<-data.frame(t=t, y=y)
> plot(t,y)
> lstartbad<-c(a=1, b=1, c=1)
> lstart2<-c(a=100, b=10, c=1)
> nlsr::nlxb(lform, data=ldata, start=lstart2)
nlsr object: x 
residual sumsquares =  1.007e-19  on  21 observations
    after  13    Jacobian and  19 function evaluations
  name            coeff          SE       tstat      pval      gradient    JSingval   
a                    100     2.679e-11  3.732e+12  1.863e-216  -6.425e-11       626.6  
b                    0.1      3.78e-13  2.646e+11  9.125e-196  -3.393e-08       112.3  
c                   0.51       6.9e-13  7.391e+11  8.494e-204   1.503e-08       2.791  
> nls(lform, data=ldata, start=lstart2, trace=TRUE)
40346.    (1.08e+00): par = (100 10 1)
11622.    (2.93e+00): par = (101.47 0.49449 0.71685)
5638.0    (1.08e+01): par = (102.23 0.38062 0.52792)
642.08    (1.04e+01): par = (102.16 0.22422 0.41935)
97.712    (1.79e+01): par = (100.7 0.14774 0.45239)
22.250    (1.78e+02): par = (99.803 0.093868 0.50492)
0.025789  (1.33e+03): par = (100.01 0.10017 0.50916)
6.0571e-08 (7.96e+05): par = (100 0.1 0.51)
4.7017e-19 (1.86e+04): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
... (approx 40 lines omitted)
1.2440e-27 (5.71e-01): par = (100 0.1 0.51)
Error in nls(lform, data = ldata, start = lstart2, trace = TRUE) : 
  number of iterations exceeded maximum of 50
> # Fix with scaleOffset
> nls(lform, data=ldata, start=lstart2, trace=TRUE, control=list(scaleOffset=1.0))
40346.      (1.08e+00): par = (100 10 1)
11622.      (2.91e+00): par = (101.47 0.49449 0.71685)
5638.0      (9.23e+00): par = (102.23 0.38062 0.52792)
642.08      (5.17e+00): par = (102.16 0.22422 0.41935)
97.712      (2.31e+00): par = (100.7 0.14774 0.45239)
22.250      (1.11e+00): par = (99.803 0.093868 0.50492)
0.025789    (3.79e-02): par = (100.01 0.10017 0.50916)
6.0571e-08  (5.80e-05): par = (100 0.1 0.51)
4.7017e-19  (1.62e-10): par = (100 0.1 0.51)
Nonlinear regression model
  model: y ~ a/(1 + b * exp(-c * t))
   data: ldata
     a      b      c 
100.00   0.10   0.51 
 residual sum-of-squares: 4.7e-19
Number of iterations to convergence: 8 
Achieved convergence tolerance: 1.62e-10

Running under: Linux Mint 20.3
```

@@

The single convergence criterion of `nls()` leaves out some possibilities that could
be useful for some problems. The package `nlsr` (@nlsr-manual) already offers both 
the safeguarded relative offset test (**roffset**) as well as a **small sum of 
squares** test (**smallsstest**) that compares the latest evaluated sum of squared
(weighted) residuals to `e4` times the initial sum of squares. The value of 
`e4` is computed as

@@

We note that `nls()` uses a termination test to stop after `maxiter` "iterations". 
Unfortunately, for almost all
iterative algorithms, the meaning of "iteration" requires careful examination of the
code. we prefer to express such tests using the number of times the residuals or the jacobian have
been computed and put upper limits on these. Our codes exit (terminate) when these limits
are reached. Generally we prefer larger limits than the default `maxiter=50` of `nls()`,
but that may simply reflect our history of dealing with more difficult problems as we are
the tool-makers; users consult us when things go wrong.

@@

gets over 4000 hits that are spread over the years. 
In some cases this is due
to the failure of the simple finite difference approximation of the Jacobian in the
`numericDeriv()` function that is a part of `nls()`. `nlsr` can use analytic
derivatives, and we could import this functionality to the `nls()` code as an
improvement. See below in the section **Jacobian computation**.

A common source of the issue is that the Jacobian is very close
to singular for some values of the model parameters. In such cases we need to find
an alternative algorithm to the Gauss-Newton iteration of `nls()`. The most common
work-around is the Levenberg-Marquardt stabilization (see @Marquardt1963, @Levenberg1944, 
@jn77ima). Versions of this have been implemented in packages `minpack.lm` and `nlsr`.
and we have prepared experimental versions of `nls` replacements that can incorporate
stabilizations. Integration of such ideas with other code structures so that all the
features of `nls()` work properly has unfortunately proved difficult.

Let us consider an example which presents the infamous 'singular gradient' error message.
This was posted on StackExchange. 
*https://stats.stackexchange.com/questions/13053/singular-gradient-error-in-nls-with-correct-starting-values*

```{r SESingGrad}
reala=-3
realb=5
realc=0.5
realr=0.7
realm=1
x=1:11 #x values 11 timepoint data
#linear+exponential function
y=reala + realb*realr^(x-realm) + realc*x
testdat=data.frame(x,y)
strt<-list(a=-3, b=5, c=0.5, r=0.7, m=1)
jform<-y~a+b*r^(x-m)+c*x # Formula
linexp=try(nls(jform, data=testdat, start=strt, trace=F))
library(nlsr)
# Note singular values of Jacobian in rightmost column
linexp2<-try(nlxb(jform, data=testdat, start=strt, trace=F))
linexp2
```

In the above, we see that the Jacobian is essentially singular as shown by its singular
values. Note that we DISPLAY the singular values in a single column for convenience in
the output. The values do NOT correspond to individual parameters, but are a property
of the whole problem.

@@

`nls()`, with the `numericDeriv()` function, computes the Jacobian as the "gradient"
attribute of the residual vector. This is implemented as a mix of R and C code, but
we have created a rather more compact version entirely in R.
Code project. See the companion document **Jacobian Calculations for nls()**
(?? need citation -- currently **DerivsNLS.pdf**).

The Jacobian used by `nlsr::nlxb()` is computed from analytic expressions, while that
of `nls()` will use a numerical approximation. In the most recent example above,
the singularity is not due to the numerical approximation.

It should be noted that the selfStart models in the `./src/library/stats/R/zzModels.R` file
provide the Jacobian in the "gradient" attribute of the "one-sided" formula that defines each
model, and these Jacobians may be the analytic forms. The `nls()` function, after computing the
"right hand side" or `rhs` of the residual, then checks to see if the "gradient" attribute is
defined, and, if not, uses `numericDeriv` to compute a Jacobian into that attribute. This code
is within the `nlsModel()` or `nlsModel.pliner()` functions. The use of analytic Jacobians
almost certainly contributes to the performance of `nls()` on selfStart models.

@@

While the implementation of subset at the level of the call to `nls()` has a certain
attractiveness, it does mean that the programmer of the solver needs to be aware of the
source (and value) of objects such as the data, residuals and Jacobian. By preference, 
we would implement subsetting by means of zero-value weights, with observation counts
(and degrees of freedom) computed via the numbers of non-zero weights. Alternatively,
we would extract a working dataframe from the relevant elements in the original.

@@

A succinct description of the issue is given in: *https://stats.stackexchange.com/questions/492955/should-i-use-na-omit-or-na-exclude-in-a-linear-model-in-r*
